---
title: "GMS_SVF"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE)
```

## Loading Library
```{r library}
library(raster)
library(horizon)
library(rgdal)
library(data.table)
pro=CRS("+init=epsg:28992")
WGS84<-CRS("+init=epsg:4326")
```

### GMS locations
[see](http://www.ahn.nl/pagina/open-data.html) for the ahn_units. 
```{r}

ahn.units<-readOGR(dsn="AHN_kaartbladen",layer="ahn_units")
proj4string(ahn.units)<-pro
plot(ahn.units)
coords.gms<-fread("data/GMSlocatie_marcel_jan2016.csv")
coords.gms<-na.omit(coords.gms)
coords.gms<-data.frame(coords.gms)
coordinates(coords.gms)<-~loc_lon+loc_lat
proj4string(coords.gms)<-WGS84
coords.gms<-spTransform(coords.gms,pro)


ov<-over(coords.gms,ahn.units)
combined<-cbind(ov,as.data.frame(coords.gms))
# saveRDS(combined,"data/gms_ahn2.rda")

```

### For station the bilt
```{r}
i=1
station<-combined[i,]
unit<-as.character(station$UNIT)

```

> HERE COMES CODE TO LOAD THE UNIT OF THE AHN2
OR SOME CODE TO LOAD THE LAZ FROM AHN3 (IF AVAILABLE)

the coding is something like (for AHN3):
unit=01cz1
file=C_01CZ1.LAZ

while AHN2 objects folder splits the files in:
u01cz1_(number 1:25).laz

for the terrain a code like:
q01cz(number 1:25).laz

The download link with the kaartblad indeling is no longer available, so I didnt' find an alternative to download subunits, which we need in the case of AHN2.

```{r}
r.Bilt<-raster("AHN/r32cn1.tif") #runnning the 0.5m tif takes approximately 2 hours
plot(r.Bilt,main="Altitude map of De Bilt")
points(coords.gms,col="red",lwd=3)
print(r.Bilt)
r.Bilt<-aggregate(r.Bilt,fact=10) #regridding to a 5m grid, this takes just a minute to run
res(r.Bilt)
plot(r.Bilt)
points(coords.gms,col="red",lwd=3)
system.time(
s.Bilt <- svf(r.Bilt, nAngles=16, maxDist=100, ll=F)
)

plot(s.Bilt,main="SVF for De Bilt")
points(coords.gms,col="red",lwd=1)
```
