---
title: "Calculating the Sky View Factor using the horizon package"
author: "Marieke"
date: "November 21, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE)
```

## Loading Library
```{r library}
library(raster)
library(horizon)
library(rgdal)
pro=CRS("+init=epsg:28992")
WGS84<-CRS("+init=epsg:4326")
```

## Loading the data
The altitude data has a resolution of 90 meters. 
```{r loading austria}
#r <- getData('alt', country='Austria') #used for downloading the data
r <- raster("data/AUT_msk_alt.grd")
plot(r,main="Altitude map of Austria")
print(r)

```

## Loading the data
```{r loading netherlands}
#r.NED <- getData('alt', country='NL')# used for downloading the data
r.NED <- raster("data/NLD_msk_alt.grd")
plot(r.NED,main="Altitude map of Netherlands")
print(r.NED)

```

## Calculating the SVF

```{r calculations}
s <- svf(r, nAngles=16, maxDist=500, ll=TRUE)
s.NED <- svf(r.NED, nAngles=16, maxDist=500, ll=TRUE)
print(s)
print(s.NED)
```

## Plotting routine

```{r plotting}
plot(s,main="SVF for Austria")
plot(s.NED,main="SVF for Netherlands")
```

## A higher resolution example: AHN
The raw data from the AHN is available with this [link](http://geodata.nationaalgeoregister.nl/ahn2/atom/ahn2_05m_ruw.xml). As an example I included the map number r50fn2.

Besides this product from PDOK and nationaal Georegister also a 3D map of the Netherlands is [available](https://www.pdok.nl/nl/producten/pdok-downloads/basisregistratie-topografie/3d-kaart-nl)

```{r}
xml.link.raw<-"http://geodata.nationaalgeoregister.nl/ahn2/atom/ahn2_05m_ruw.xml" 
r.NED<-raster("AHN/r09dn1.tif") #runnning the 0.5m tif takes approximately 2 hours
plot(r.NED,main="Altitude map of Netherlands")
print(r.NED)
```

### Regridding the data
The runtime for a 0.5m resolution grid is long. Therefore I regrid the raster to a 5m grid. Calculating the SVF will now take approximately 1 minute.
```{r}
r.NED<-aggregate(r.NED,fact=10) #regridding to a 5m grid, this takes just a minute to run
res(r.NED)
plot(r.NED)
system.time(
s.NED <- svf(r.NED, nAngles=16, maxDist=100, ll=F)
)
plot(s.NED,main="SVF for Netherlands")
```

### Other area's
[see](http://www.ahn.nl/pagina/open-data.html) for the ahn_units. 
```{r}


ahn.units<-readOGR(dsn="AHN_kaartbladen",layer="ahn_units")
proj4string(ahn.units)<-pro
plot(ahn.units)
coords.knmi<-readRDS("data/coordsKNMI.rda")
coords.knmi<-na.omit(coords.knmi)
coords.knmi<-data.frame(coords.knmi)
coordinates(coords.knmi)<-~DS_LON+DS_LAT
proj4string(coords.knmi)<-WGS84
coords.knmi<-spTransform(coords.knmi,pro)


ov<-over(coords.knmi,ahn.units)
combined<-cbind(ov,as.data.frame(coords.knmi))
print(combined[which(combined$DS_CODE=="260_H"),])
```

### For station the bilt
```{r}
r.Bilt<-raster("AHN/r32cn1.tif") #runnning the 0.5m tif takes approximately 2 hours
plot(r.Bilt,main="Altitude map of De Bilt")
points(coords.knmi,col="red",lwd=3)
print(r.Bilt)
r.Bilt<-aggregate(r.Bilt,fact=10) #regridding to a 5m grid, this takes just a minute to run
res(r.Bilt)
plot(r.Bilt)
points(coords.knmi,col="red",lwd=3)
system.time(
s.Bilt <- svf(r.Bilt, nAngles=16, maxDist=100, ll=F)
)

plot(s.Bilt,main="SVF for De Bilt")
points(coords.knmi,col="red",lwd=1)
```