---
title: "Calculating the Sky View Factor using the horizon package"
author: "Marieke"
date: "November 21, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE)
```

## Loading Library
```{r library}
library(raster)
library(horizon)
library(rgdal)
pro=CRS("+init=epsg:28992")
WGS84<-CRS("+init=epsg:4326")
```

## Loading the data
The altitude data has a resolution of 90 meters. 
```{r loading austria}
#r <- getData('alt', country='Austria') #used for downloading the data
r <- raster("data/AUT_msk_alt.grd")
plot(r,main="Altitude map of Austria")
print(r)

```

## Loading the data
```{r loading netherlands}
#r.NED <- getData('alt', country='NL')# used for downloading the data
r.NED <- raster("data/NLD_msk_alt.grd")
plot(r.NED,main="Altitude map of Netherlands")
print(r.NED)

```

## Calculating the SVF

```{r calculations}
s <- svf(r, nAngles=16, maxDist=500, ll=TRUE)
s.NED <- svf(r.NED, nAngles=16, maxDist=500, ll=TRUE)
print(s)
print(s.NED)
```

## Plotting routine

```{r plotting}
plot(s,main="SVF for Austria")
plot(s.NED,main="SVF for Netherlands")
```

## A higher resolution example: AHN
The raw data from the AHN is available with this [link](http://geodata.nationaalgeoregister.nl/ahn2/atom/ahn2_05m_ruw.xml). As an example I included the map number r50fn2.

Besides this product from PDOK and nationaal Georegister also a 3D map of the Netherlands is [available](https://www.pdok.nl/nl/producten/pdok-downloads/basisregistratie-topografie/3d-kaart-nl)

```{r}
xml.link.raw<-"http://geodata.nationaalgeoregister.nl/ahn2/atom/ahn2_05m_ruw.xml" 
r.NED<-raster("AHN/r09dn1.tif") #runnning the 0.5m tif takes approximately 2 hours
plot(r.NED,main="Altitude map of Netherlands")
print(r.NED)
```

### Regridding the data
The runtime for a 0.5m resolution grid is long. Therefore I regrid the raster to a 5m grid. Calculating the SVF will now take approximately 1 minute.
```{r}
r.NED<-aggregate(r.NED,fact=10) #regridding to a 5m grid, this takes just a minute to run
res(r.NED)
plot(r.NED)
system.time(
s.NED <- svf(r.NED, nAngles=16, maxDist=100, ll=F)
)
plot(s.NED,main="SVF for Netherlands")
```

### Other area's
[see](http://www.ahn.nl/pagina/open-data.html) for the ahn_units. 
```{r}


ahn.units<-readOGR(dsn="AHN_kaartbladen",layer="ahn_units")
proj4string(ahn.units)<-pro
plot(ahn.units)
coords.knmi<-readRDS("data/coordsKNMI.rda")
coords.knmi<-na.omit(coords.knmi)
coords.knmi<-data.frame(coords.knmi)
coordinates(coords.knmi)<-~DS_LON+DS_LAT
proj4string(coords.knmi)<-WGS84
coords.knmi<-spTransform(coords.knmi,pro)


ov<-over(coords.knmi,ahn.units)
combined<-cbind(ov,as.data.frame(coords.knmi))
print(combined[which(combined$DS_CODE=="260_H"),])
```

### For station the bilt
```{r}
r.Bilt<-raster("AHN/r32cn1.tif") #runnning the 0.5m tif takes approximately 2 hours
plot(r.Bilt,main="Altitude map of De Bilt")
points(coords.knmi,col="red",lwd=3)
print(r.Bilt)
r.Bilt<-aggregate(r.Bilt,fact=10) #regridding to a 5m grid, this takes just a minute to run
res(r.Bilt)
plot(r.Bilt)
points(coords.knmi,col="red",lwd=3)
system.time(
s.Bilt <- svf(r.Bilt, nAngles=16, maxDist=100, ll=F)
)

plot(s.Bilt,main="SVF for De Bilt")
points(coords.knmi,col="red",lwd=1)
```

### 3D map of the Netherlands: A combined product, top10NL and AHN2
A high resolution map of the Netherlands is [available](http://www.arcgis.com/apps/Viewer/index.html?appid=2e4c2b5b127a447e94997bbffd41b93e) and with 97% almost complete. More background information can be found in the [technical description](http://www.kadaster.nl/documents/20838/88032/3D+kaart+NL+technische+beschrijving/841392d9-12e2-41f8-b20d-32d73fbfbbf9 ) of the product. The link to a code at the end of the document provides more information on reading the data using the comand prompt. The [code](http://data.nlextract.nl/top10nl3d/README.txt) takes approximately 1.5hour per unit to run. An example for Valkenburg is [included](https://3d.bk.tudelft.nl/hledoux/blog/tinkering-3dtop10nl/). 

The structure of the data is not so trivial that you can simply read it in R. You first need to know the structure of the database and the layer names before using the readOGR command. An online example can be found [here](http://gis.stackexchange.com/questions/151613/how-to-read-feature-class-in-file-geodatabase-using-r). However, I did not yet find a way to solve the following warning:

> Warning message: In readOGR(dsn = file.gbd, layer = "terreinVlak_3D_LOD0") : Z-dimension discarded

```{r top10NL De Bilt}
file.gbd<-"AHN/top10NL/32cn1.gdb"
subset(ogrDrivers(), grepl("GDB", name))
fc_list = ogrListLayers(file.gbd)
print(fc_list)

buildings<-readOGR(dsn=file.gbd,layer="gebouwVlak")
terrein<-readOGR(dsn=file.gbd,layer="terreinVlak")
water<-readOGR(dsn=file.gbd,layer="waterdeelVlak")

plot(terrein)
plot(water,col="yellow",add=TRUE)
```
